### 2025-04-05 basic-statistics.qmd
20250405 - descriptive stats

exampleValues <- c(0, 0, 10, 15, 20, 30, 1000)
exampleValues_mean <- apa(mean(exampleValues), 2)
```

That is, to compute the mean, sum all of the values and divide by the number of values ($n$).
One issue with the mean is that it is sensitive to extreme (outlying) values.
For instance, the mean of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mean`.

@@

The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
The pseudomedian can be computed using the `DescTools` package [@R-DescTools].
The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.

@@

```{r}
mean(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
median(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
DescTools::HodgesLehmann(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
petersenlab::Mode(player_stats_seasonal$fantasyPoints)

weighted.mean(
  player_stats_seasonal$fantasyPoints,
  weights = sample( # randomly generate weights (could specify them manually)
    x = 1:3,
    size = length(player_stats_seasonal$fantasyPoints),
    replace = TRUE),
  na.rm = TRUE)

petersenlab::wthdquantile(
  player_stats_seasonal$fantasyPoints,
  w = sample( # randomly generate weights (could specify them manually)
    x = 1:3,
    size = length(player_stats_seasonal$fantasyPoints),
    replace = TRUE),
  probs = 0.5)
```

@@

```{r}

```

@@

```{r}

```

@@

player_stats_seasonal %>% 
  select(age, years_of_experience, fantasyPoints) %>% 
  summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    pivot_wider(
      names_from = index,
      values_from = value)
```

>>>>

exampleValues <- c(0, 0, 10, 15, 20, 30, 1000)
exampleValues_mean <- mean(exampleValues)
```

That is, to compute the mean, sum all of the values and divide by the number of values ($n$).
One issue with the mean is that it is sensitive to extreme (outlying) values.
For instance, the mean of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} petersenlab::apa(exampleValues_mean, decimals = 2)`.

@@

The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
The pseudomedian can be computed using the `DescTools` package [@R-DescTools].
The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.
The pseudomedian is an example of a robust average—an index of the center of a distribution that is less sensitive to outliers.

@@

- mean:

```{r}
mean(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- median:

```{r}
median(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- Hodges-Lehmann statistic (aka pseudomedian):

```{r}
DescTools::HodgesLehmann(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- mode:

```{r}
petersenlab::Mode(player_stats_seasonal$fantasyPoints)
```

- weighted mean:

```{r}
weighted.mean(
  player_stats_seasonal$fantasyPoints,
  weights = sample( # randomly generate weights (could specify them manually)
    x = 1:3,
    size = length(player_stats_seasonal$fantasyPoints),
    replace = TRUE),
  na.rm = TRUE)
```

- weighted median:

```{r}
petersenlab::wthdquantile(
  player_stats_seasonal$fantasyPoints,
  w = sample( # randomly generate weights (could specify them manually)
    x = 1:3,
    size = length(player_stats_seasonal$fantasyPoints),
    replace = TRUE),
  probs = 0.5)
```

@@

- standard deviation:

```{r}
sd(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- variance:

```{r}
var(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- range:

```{r}
range(player_stats_seasonal$fantasyPoints, na.rm = TRUE)

max(player_stats_seasonal$fantasyPoints, na.rm = TRUE) - min(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- minimum and maximum:

```{r}
min(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
max(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- interquartile range (IQR):

```{r}
IQR(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- median absolute deviation:

```{r}
mad(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

@@

- skewness:

```{r}
psych::skew(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

- kurtosis:

```{r}
psych::kurtosi(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
```

@@

player_stats_seasonal %>% 
  dplyr::select(age, years_of_experience, fantasyPoints) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```
